Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-19 2017-Xudong-SMT/FM-2018-extended.tex
Added justification of properties in Section 5.4; Proof-reading, reduction and notes in other sections.

\paragraph{Related works}

@@

There is not much research work towards trying to develop symbolic bisimulation approaches
for the value-passing process algebra and languages, as our long-term goals,
especially no algorithm treatment of the symbolic systems developed by
interacting with automatic theorem provers. The closest work is the
one already mentioned from J. Rathke~\cite{HennessyRathke:TCS98},
who developed the symbolic bisimulation for a
calculus of broadcasting system (CBS). CBS is similar with classic
process calculi such as CCS and CSP, but communicating by broadcasting
values, one-to-many communication instead of one-to-one communication
and transmitting values without blocking. That makes the definition of
the symbolic semantic and bisimulation equivalence different from the
classic works.

@@

\paragraph*{Term algebra.}
Our models rely on a notion of parameterized actions, that are
symbolic expressions using data types and variables. As we want to encode
the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

\def\Talg{\mathcal{T}_{\Sigma,\P}}
Formally, we assume the existence of a term algebra $\Talg$,
where $\Sigma$ is the signature of the data and action constructors,
and $\P$ a set of variables. Within $\Talg$, we distinguish a set of
data expressions $\mathcal{E}_\P$, including a set of boolean
expressions $\mathcal{B}_{\P}$ ($\mathcal{B}_{\P}\subseteq\mathcal{E}_\P$).
On top of $\mathcal{E}_\P$ we build the action algebra
$\mathcal{A}_\P$, with $\mathcal{A}_P\subseteq\mathcal{T}_\P,
\mathcal{E}_P\cap\mathcal{A}_P=\emptyset$;
naturally action terms will use data expressions as sub-terms.
The function $\vars(t)$ identifies the set of variables in a term
$t\in\AlgT$.

pNets can encode naturally the notion of input actions as found
e.g. in value-passing CCS 
\cite{Milner89} or of usual point-to-point message passing calculi,
but it also allows 
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.

\paragraph*{Algebra presentations.}
In practice, the parameterization of the pNet model by some specific
action algebra is realized by the definition of an ``algebra
presentation''. It will be used to check the
well-formness of a pNet system, and to define the translation of the pNet
semantics into the SMT engine input language.

\begin{definition}
  An Algebra Presentation is a pair $\mathcal{P}=<Sorts,Constrs,Ops>$ where:

  \begin{itemize}
  \item $Sorts$ is a set of Sort names
    \footnote{Later we may want to extend this with Sort constructors,
      like Array or Pair, but this is not needed now}
  \item $Constrs$ is a set of constructor operators, with $Con$ the
    constructor name, and arity: $arity(Con)=n \in \mathbb{N}$,
    and with their signature and their associated selectors,
    of the form: '$Con : (sel_1,sort_1), ... , (sel_n,sort_n) -> sort$'.
    For each argument, the pair $(sel_i,sort_i)$ defines an auxiliary
    operator of name $sel_i$ with signature $sel_i : sort -> sort_i$.
    \item $Ops$ is a set of other (auxiliary) operators, with their
      arity and signature, of the form: $Op : sort_1, ...  sort_n ->
      sort$
      \item $Constrs(sortname), Sels(sortname)$, respectively define the set of
        constructors and selectors of sort $sortname$
  \end{itemize}
  All sorts and operator names must be distinct.\\
  Amongst Constructors, those of arity 0 are called constants, and we
  define $Consts(\mathcal{P}) = \{Con \in Constrs. arity(Con)=0\}$.
\end{definition}

Sorts \texttt{Bool} and \texttt{Int} are predefined with standard operators.
Sort \texttt{Action} also, with a constructor \texttt{Synchro} denoting
a synchronized action, i.e. an ``internal'' action that cannot be
further synchronized with the environment. It also comes with an
overloaded \texttt{FUN} constructor, used to build actions with
arguments, that will be instanciated to the required sorts for a given
pNet.

\begin{table}\caption{\label{table:BIPalgebra}\leftline{Algebra
      Presentation: predefined Sorts and Operators}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Other Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\texttt{true},\ \texttt{false}$&
                $\land,\ \lor,\ \neg,\ \implies,\ =,\ \ne$
                \\\specialrule{0em}{1pt}{1pt} 
                %               		Action 			&  $\texttt{Synchro},\ \texttt{FUN\_Action\_Bool}$ &
                Action 			&  $\texttt{Synchro},\ \texttt{FUN}$ &
                \\\specialrule{0em}{1pt}{1pt}
		Int 				&
                ${0, \{i, -i\}_{i \in \texttt{Nat}}}$  &
                $- \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div, \texttt{etc.}$
%                \\\specialrule{0em}{1pt}{1pt}
%                \textsl{for any sort} & & $=,\ \ne$
		\\\hline\specialrule{0em}{1pt}{1pt}
\multicolumn{3}{l}{\sl Extension for the BIP use-case of Fig.~\ref{schema:ArchFailure:pNet}}
		\\\hline\specialrule{0em}{1pt}{1pt}
                Action & \multicolumn{2}{l} \texttt{FUN\_Action\_Bool}, \texttt{}fail, resume, timeout, reset, start, tick, ask 
		\\\hline
	\end{tabular}
\end{table}

@@

A pLTS is a labelled transition system with variables; variables can be
manipulated, defined, or accessed inside states, actions, guards, and
assignments.

%Without loss of generality and to simplify the formalization, we suppose 
%here that variables are local to 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment.

Each state has its set of variables called $State\ variables$, 
which can only be modified by the assignment in the transitions targeting its owner state. 
A global state variable of a pLTS is a variable $x$ satisfying $\forall s\!\in\! S .\, x\!\in\! \vars(s)$.

%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
Note that we make no assumption on finiteness of the set of states nor
on finite branching of the transition relation.

We first define the set of actions a pLTS can use, let $a$
range over action labels, $\symb{op}$ are operators, and $x_i$ range over
variable names. Action terms are:
\[
\begin{array}[l]{rcl@{\quad}p{5.5cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ~\symb{Expr}&\text{parameters}\\
  \symb{Expr}&::=& \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,..,\symb{Expr}_n)&\text{Expressions}
\end{array}
\]
%\[
%\begin{array}[l]{rcl@{\quad}p{5.5cm}}
%  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
%  p_i&::=& ?x~|~\symb{Expr}&\text{parameters (input variable or expression)}\\
%  \symb{Expr}&::=& \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,..,\symb{Expr}_n)&\text{Expressions}
%\end{array}
%\]

@@

\begin{definition}[pLTS]
\label{pLTS}
A pLTS is a tuple
$pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of states.
\item[$\bullet$]
$s_0 \in S$ is the initial state.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the transition relation and 
$L$ is the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterized action, $e_b \in
\AlgB$ is a guard, and the variables $x_j\in P$
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
		$\vars(\alpha)\!\subseteq\! \vars(s)$, 
		$\vars(e_b)\!\subseteq\! \vars(s')$, and
		$\forall j\!\in\! J .\,\vars(e_j)\!\subseteq\! \vars(s)\land 
x_j\!\in\!\vars(s')$.
\item[$\bullet$]
The $x_j$ here are $state\ variables$ of state $s'$.

@@

Now we define
pNet nodes, as constructors for hierarchical behavioral structures.
A pNet node has a set of sub-pNets that can be either pNets or pLTSs, and a
set of Holes, playing the role of process parameters.

A composite pNet consists of a set of sub-pNets exposing
a set of actions, each of them triggering internal actions in each of
the sub-pNets. The synchronization between global actions and
internal actions is given by  \emph{synchronization vectors}: a
synchronization vector synchronizes one or several internal actions, and
exposes a single resulting global action.

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure which leaves are pLTSs and holes:\\
$\pNet\triangleq pLTS~|~\mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $I \in \I$ is the set over which sub-pNets are indexed.
\item[$\bullet$] $\pNet_i^{i\in I}$ is the family of sub-pNets.

\item[$\bullet$] $J\in\I_\P$ is the set over which holes are indexed.
$I$ and $J$ are \emph{disjoint}: $I\cap J=\emptyset$,  $I\cup J\neq\emptyset$

\item[$\bullet$] $\symb{SV}_k^{k\in K}$ is a set of
  synchronization vectors ($K\in\I_\P$). $\forall k\!\in\! K,
  \symb{SV}_k\!=\!\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k$ where
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\! I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$,
  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!S_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\Label(\symb{SV}_k) = \alpha'_k$.

In Fig.~\ref{schema:ArchFailure:pNet} we show examples of these
constructs, with 2 pLTSs and one pNet node encoding our use-case.

@@

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified\,\footnote{%
%
  The current approach relies on taxonomies of predefined
  architectures.  Future work may include automatic synthesis of
  architectures from, \eg[,] CTL~\cite{BaierKatoen2008} formulae.
%
} and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

@@

Finally, \emph{priorities}\mdash defined by a strict partial order on
the set of possible interactions\mdash narrow the choice among the
enabled interactions at any given round.  The default priority is the
so-called \emph{maximal progress}, whereby among any two interactions
$a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
For example, the port $\NameOprnd.\PortFail$ will never fire alone in
a global state, where both $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are enabled.
%
Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
$\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
a failure, the operation performed by $\NameOprnd$ cannot finish
unless it is explicitly resumed or the system is reset.

@@

Failure Monitor enforces 1)~the safety property \emph{``the
  system reset never happens, unless asked for by a timeout
  following a failure''}, formalised in CTL by
%
\[
\varphi \land \mathtt{AG} (\PortReset \rightarrow \varphi)\,,
\quad\text{where}\quad
%
\varphi = \mathtt{A} [\lnot \PortReset\ \mathtt{W}\ \PortAsk]
\land \mathtt{A} [\lnot \PortAsk\ \mathtt{W}\ \PortFail]
\]
%
($\mathtt{W}$ being the \emph{weak until} operator) and 2)~the
liveness property \emph{``a reset will be fired when asked for by
  a timeout''}:
%
$\mathtt{AG}(\PortAsk \rightarrow \mathtt{AF}\ \PortReset)$.
%

\TODO{it would be good to express here maybe 2 properties of the
  architecture, maybe one taken from the original study, and another
  one involving the timer data (number of ticks ?)}

@@

In this paper we presented the algorithm for generating an open
automaton representing the semantics of an open pNet, and we described
the implementation of the
algorithm. Our implementation includes two main parts. First, we
compute all the possible open transitions. The actions in pLTS and
hole behaviors are composed in the algorithm while the generation of
predicate need to match the combinations of the subnets with the
synchronization vectors. Some of the open transitions obtained at this
intermediate step, constructed in a structural manner from all
possible combinations of possible logical predicates, may
contain predicates which do not represent any possible concrete
instantiations. 
To get rid of these useless transitions, we use the SMT solver Z3 for checking the
satisfiability of the predicate in each open transition. In
order to do that, we encode into Z3 the action algebra presentation,
the a representation of the predicates, before submitting them to the
Z3 solver. We implemented our algorithm in the VerCors platform and use some
running examples, encoding expressions from various process algebras, to test the
algorithm. Our use-cases show that our encoding
identifies successfully all unsatisfiable open transitions and that
the resulting automaton reflects correctly the expected 
movements of the encoded process expressions.

Among the questions arising during this work, an important one is the
efficiency of Z3 for our needs. Z3 is famous for being very fast for
solving very large sets of assertions, and this could be important for
us. But we encountered  some problems during the implementation the
checking method using Z3. For example, if the $functions$ and the
$datatypes$ part of the algebra presentation we submit to the solver
are independent with each other, then checking the $assertions$ is a
simple work. However, $functions$ on recursive $datatypes$ make it
more complex, some special rules might be defined by user for the
induction. Also, we need more evidence that dealing with more complex
action algebras, that would involve axiomatizing their data structures
into Z3 theories, will indeed allow us to decide validity of
predicates.
We may also try other automatic theorem provers, depending on the structure of the proofs we need.

@@

Naturally, our next goals after the generation of the open automata
will be to model-check, and to check equivalence of  pNets. While
model-checking open automata (with a suitable logic including
predicates on data) seems easy to define, equivalence checking is more challenging.
In previous paper, we
have already found the FH-bisimulation, inspired by~\cite{deSimone85} as a
suitable definition. But usual approaches of bisimulation by partition
refinement does not work easily with symbolic transitions. In a first
step we may want instead to define and implement an algorithm for
checking that a given equivalence relation between open automata 
states is indeed a FH-Bisimulation. Finding automatically a
bisimulation relation or the most general bisimulation will be more
challenging. It appears also that when comparing different pNets with
different structure, strong bisimulation will be too strong a notion,
and that we will have to define proper notions of weak equivalence or
refinements.

@@

\begin{table}\caption{\label{table:CCSalgebra}\leftline{Additional
      sorts and operators for the CCS algebra}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Other Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Channel, Data    			&
                $\emptyset$&
                $\emptyset$
                \\\specialrule{0em}{1pt}{1pt} 
		Action 			&  $\texttt{Emit(c,v)},\texttt{Receive(c,v)},\texttt{Tau}$ &
		\\\hline
	\end{tabular}
\end{table}

@@

		&&       		\\\specialrule{0em}{1pt}{1pt}
		\multicolumn{2}{l}{Other operators: }

>>>>

\paragraph{Related work.}

@@

%% There is not much research work towards trying 
Very few attempts were made to develop symbolic bisimulation approaches
for the value-passing process algebras and languages\mdash our long-term goals\mdash
especially, there is no algorithmic treatment of the symbolic systems developed by
interacting with automatic theorem provers. The closest work is the
one already mentioned from J. Rathke~\cite{HennessyRathke:TCS98},
who developed the symbolic bisimulation for a
calculus of broadcasting system (CBS). CBS is similar with classic
process calculi such as CCS and CSP, but communicating by broadcasting
values, one-to-many communication instead of one-to-one communication
and transmitting values without blocking. That makes the definition of
the symbolic semantic and bisimulation equivalence different from the
classic works.

@@

\paragraph*{Term algebra.}
Our models rely on a notion of parameterized actions that are
symbolic expressions using data types and variables. As we want to encode
the low-level behavior of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on unification between two or more action
expressions. This will be used in the semantics of synchronization
vectors to express various kinds of communication or synchronization mechanisms.

\def\Talg{\mathcal{T}_{\Sigma,\P}}
Formally, we assume the existence of a term algebra $\Talg$,
where $\Sigma$ is the signature of the data and action constructors,
and $\P$ a set of variables. Within $\Talg$, we distinguish a set of
data expressions $\mathcal{E}_\P$, including a set of Boolean
expressions $\mathcal{B}_{\P}$ ($\mathcal{B}_{\P}\subseteq\mathcal{E}_\P$).
On top of $\mathcal{E}_\P$ we build the action algebra
$\mathcal{A}_\P$, with $\mathcal{A}_P\subseteq\mathcal{T}_\P,
\mathcal{E}_P\cap\mathcal{A}_P=\emptyset$;
naturally action terms will use data expressions as sub-terms.
The function $\vars(t)$ identifies the set of variables in a term
$t\in\AlgT$.

pNets can encode naturally the notion of input actions as found,
\eg in value-passing CCS 
\cite{Milner89} or of usual point-to-point message passing calculi,
but it also allows 
for more general mechanisms, like gate negotiation in Lotos~\hl{\cite{lotos}}, or broadcast
communications.

\paragraph*{Algebra presentations.}
In practice, the parameterization of the pNet model by some specific
action algebra is realized by the definition of an ``algebra
presentation''. It will be used to check the
well-formedness of a pNet system, and to define the translation of the pNet
semantics into the SMT engine input language.

\begin{definition}
  An \emph{algebra presentation} is a triple $\mathcal{P}=\langle\mathit{Sorts},\mathit{Constrs},\mathit{Ops}\rangle$, where

  \begin{itemize}
  \item $Sorts$ is a set of \emph{sorts}\noteSB{Commented out a footnote here\mdash check if still necessary.}
    %% \footnote{Later we may want to extend this with Sort constructors,
    %%   like Array or Pair, but this is not needed now}
  \item $\mathit{Constrs}$ is a set of \emph{constructor operators}: for each $\mathit{Con} \in \mathit{Constrs}$, $arity(Con)=n \in \mathbb{N}$ is its arity 
    and '$Con : (\mathit{sel}_1,\mathit{sort}_1), \dots, (\mathit{sel}_n,\mathit{sort}_n) \rightarrow \mathit{sort}$' is its signature with the associated selectors.
    For each argument, the pair $(\mathit{sel}_i,\mathit{sort}_i)$ defines an auxiliary
    operator of name $\mathit{sel}_i$ with signature $\mathit{sel}_i : \mathit{sort} \rightarrow \mathit{sort}_i$.
    \item $Ops$ is a set of \emph{auxiliary operators}, with their
      arity and signature, of the form: $Op : \mathit{sort}_1, \dots,  \mathit{sort}_n \rightarrow
      \mathit{sort}$
      \item $\mathit{Constrs}(\mathit{sortname})$ and $\mathit{Sels}(\mathit{sortname})$ are, respectively, the sets of
        constructors and selectors of the sort $\mathit{sortname}$
  \end{itemize}
  \noteSB{I think this can be removed}\hl{All sorts and operator names must be distinct.}
  Among constructors, those of arity 0 are called \emph{constants}:
  $\mathit{Consts}(\mathcal{P}) = \{\mathit{Con} \in \mathit{Constrs}\,|\, \mathit{arity}(\mathit{Con})=0\}$.
\end{definition}

Sorts \texttt{Bool} and \texttt{Int} are predefined with standard operators.
Sort \texttt{Action} also, with a constructor \texttt{Synchro} denoting
a synchronized action, i.e. an ``internal'' action that cannot be
further synchronized with the environment. It also comes with an
overloaded \texttt{FUN} constructor, used to build actions with
arguments, that will be instantiated to the required sorts for a given
pNet.

\begin{table}[t]\caption{\label{table:BIPalgebra}\leftline{Algebra
      Presentation: predefined Sorts and Operators}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Auxiliary Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\texttt{true},\ \texttt{false}$&
                $\land,\ \lor,\ \neg,\ \implies,\ =,\ \ne$
                \\\specialrule{0em}{1pt}{1pt} 
                %               		Action 			&  $\texttt{Synchro},\ \texttt{FUN\_Action\_Bool}$ &
                Action 			&  $\texttt{Synchro},\ \texttt{FUN}$ &
                \\\specialrule{0em}{1pt}{1pt}
		Int 				&
                ${0, \{i, -i\}_{i \in \texttt{Nat}}}$  &
                $- \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div \text{ \etc}$
%                \\\specialrule{0em}{1pt}{1pt}
%                \textsl{for any sort} & & $=,\ \ne$
		\\\hline\specialrule{0em}{1pt}{1pt}
\multicolumn{3}{l}{\sl Extension for the BIP use-case of Fig.~\ref{schema:ArchFailure:pNet}}
		\\\hline\specialrule{0em}{1pt}{1pt}
                Action & \multicolumn{2}{l} \texttt{FUN\_Action\_Bool}, \texttt{}fail, resume, timeout, reset, start, tick, ask 
		\\\hline
	\end{tabular}
\end{table}

@@

A pLTS is a labelled transition system with variables, which can be
manipulated, defined, or accessed inside states, actions, guards, and
assignments. 
%
%Without loss of generality and to simplify the formalization, we suppose 
%here that variables are local to 
%state: each state has its set of variables disjoint from the others. Transmitting 
%variable values from one state to the other can be done by explicit assignment. 
%
Each state has its set of variables called \emph{state variables}, 
which can only be modified by the assignment in transitions targeting this state. 
A global state variable of a pLTS is a variable $x$ satisfying $\forall s\!\in\! S .\, x\!\in\! \vars(s)$.
%
%Similarly, to simplify the management of variables and without loss of expressivity, we 
%suppose that transitions looping to the same state does not do assignments.
Note that we make no assumptions on finiteness of the set of states, nor
on finite branching of the transition relation.

We first define the set of actions a pLTS can use.  Let $a$
range over action labels, $\symb{op}$ are operators, and $x_i$ range over
variable names. Action terms are:
\[
\begin{array}[l]{rcl@{\quad}p{5.5cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ~\symb{Expr}&\text{parameters}\\
  \symb{Expr}&::=& \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,\dots,\symb{Expr}_n)&\text{expressions}
\end{array}
\]
%\[
%\begin{array}[l]{rcl@{\quad}p{5.5cm}}
%  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
%  p_i&::=& ?x~|~\symb{Expr}&\text{parameters (input variable or expression)}\\
%  \symb{Expr}&::=& \symb{Value}~|~x~|~\symb{op}(\symb{Expr}_1,..,\symb{Expr}_n)&\text{Expressions}
%\end{array}
%\]

@@

\begin{definition}[pLTS]
\label{pLTS}
\noteSB{It's a bit strange that algebra presentations are not referred to here}
A pLTS is a tuple
$pLTS\triangleq\mylangle S,s_0, \to\myrangle$ where:
\begin{itemize}
\item[$\bullet$]
$S$ is a set of \emph{states}.
\item[$\bullet$]
$s_0 \in S$ is the \emph{initial state}.
\item[$\bullet$] $\to \subseteq S \times L \times S$ is the \emph{transition relation}, with
$L$ the set of labels of the form
$\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in J}\rangle$,
where $\alpha \in\AlgA$ is a parameterized action, \noteSB{Shouldn't this be $e_b \in \AlgB_{\mathcal{P}}$, for {\em some} $\mathcal{P}$?}\hl{$e_b \in
\AlgB$} is a guard, and the variables $x_j\in P$
are assigned the expressions $e_j\in \AlgE$.
If 
$s \xrightarrow{\langle \alpha,~e_b,~(x_j\!:= {e}_j)^{j\in
		J}\rangle} s'\in \to $ then 
		$\vars(\alpha)\!\subseteq\! \vars(s)$, 
		\noteSB{I would expect $\vars(e_b)\!\subseteq\! \vars(s)$, since the guard has to be tested on the source state}\hl{$\vars(e_b)\!\subseteq\! \vars(s')$}, and
		$\forall j\!\in\! J .\,\vars(e_j)\!\subseteq\! \vars(s)\land 
x_j\!\in\!\vars(s')$.
%\item[$\bullet$]
The $x_j$ are \emph{state variables} of %% the state
$s'$. 
\end{itemize}
\end{definition}

Now, we define
pNet nodes as constructors for hierarchical behavioral structures.
A pNet node has a set of sub-pNets that can be either pNets or pLTSs, and a
set of \emph{holes}, playing the role of process parameters.

A composite pNet consists of a set of sub-pNets, each exposing
a set of actions. %% , each of them triggering internal actions in each of
%% the sub-pNets.
The relation between actions of a pNet and those of its sub-pNets
%% internal actions is
are given by  \emph{synchronization vectors}, which %% : a
%% synchronization vector 
synchronize one or several internal actions, and
expose a single resulting global action.

\begin{definition}[pNets]\label{def-pnets}
A pNet is a hierarchical structure, whereof the leaves are pLTSs and holes:
$\pNet::= pLTS~|~\mylangle \pNet_i^{i\in I}, S_j^{j\in J}, \symb{SV}_k^{k\in K}\myrangle$,
where
\begin{itemize}
\item[$\bullet$] \noteSB{Why is one $\I$ and the other $\I_\P$?}\hl{$I \in \I$ and $J\in\I_\P$} are the sets over which are indexed sub-pNets and holes, respectively, such that: $I\cap J=\emptyset$ and  $I\cup J\neq\emptyset$.

\item[$\bullet$] $\pNet_i^{i\in I}$ is a family of sub-pNets.

\item[$\bullet$] $\symb{SV}_k^{k\in K}$, with $K\in\I_\P$, is a set of
  synchronization vectors: for each $k\!\in\! K$,
  $\symb{SV}_k\!=\!\alpha_{l}^{l\in I_k \uplus J_k}\to\alpha'_k$ where
  $\alpha'_k\in \mathcal{A}_\P$, $I_k\subseteq I$, $J_k\subseteq J$,
  $\forall i\!\in\! I_k.\,\alpha_{i}\!\in\!\Sort(\pNet_i)$\noteSB{What is the sort of a pNet? Should we define $\Sort(\pNet_i) \triangleq \{\Label(\symb{SV}_k\,|\,k \in K\}$?}, 
  $\forall j\!\in\!
  J_k.\,\alpha_{j}\!\in\!S_j$, and $\vars(\alpha'_k)\subseteq \bigcup_{l\in I_k\uplus 
  J_k}{\vars({\alpha_l})}$. The global action of a vector $\symb{SV}_k$ is
$\Label(\symb{SV}_k) = \alpha'_k$.

@@

In Fig.~\ref{schema:ArchFailure:pNet}, we show examples of these
constructs, with two pLTSs, one hole and one pNet node encoding our running example.

@@

In this context, the notion of \emph{architecture} was proposed
in~\cite{AttieBBJS16-architectures-faoc} as a mechanism for ensuring
correctness by construction during the design of BIP models.
Architectures can be viewed as operators transforming BIP models.
They formalise design patterns, which enforce global properties
characterising the coordination among the components of the system.
The architecture-based design process in BIP takes as input a set of
components providing basic functionality of the system and a set of
temporal properties that must be enforced in the final system.  For
each property, a corresponding architecture is identified\,\footnote{%
%
  The current approach relies on taxonomies of predefined
  architectures.  Future work may include automatic synthesis of
  architectures from, \eg CTL~\cite{BaierKatoen2008} formulae.
%
} and applied to the model, thereby potentially introducing additional
coordinator components and modifying the connectors that define
synchronisation patterns among ports of components.
In~\cite{AttieBBJS16-architectures-faoc}, it was shown that
application of architectures is compositional \wrt safety properties,
\ie when several architectures are applied, each enforcing a safety
property, the resulting system satisfies their conjunction.

@@

Finally, \emph{priorities}\mdash defined by a strict partial order on
the set of possible interactions\mdash narrow the choice among the
enabled interactions at any given round.  The default priority is the
so-called \emph{maximal progress}, whereby among any two interactions
$a \subset b$ (as sets of ports), $b$ has higher priority than $a$.
For example, the port $\NameOprnd.\PortFail$ will never fire alone in
a global state, where both $\NameTimer.\PortStart$ and
$\NameCtrl.\PortFail$ are enabled.
%
%% Adding priorities $\NameOprnd.\PortFinish < \NameCtrl.\PortReset$ and
%% $\NameOprnd.\PortFinish < \NameOprnd.\PortResume$, ensures that, after
%% a failure, the operation performed by $\NameOprnd$ cannot finish
%% unless it is explicitly resumed or the system is reset.

@@

Failure Monitor enforces 1)~the safety property \emph{``the
  system reset never happens, unless asked for by a timeout
  following a failure''}, formalised in CTL by
%
\[
\varphi \land \mathtt{AG} (\PortReset \rightarrow \varphi),
\quad\text{where}\
%
\varphi = \mathtt{A} [\lnot \PortReset\ \mathtt{W}\ \PortAsk]
\land \mathtt{A} [\lnot \PortAsk\ \mathtt{W}\ \PortFail],
\]
%
($\mathtt{W}$ being the \emph{weak until} operator) and 2)~the
liveness property \emph{``a reset will be fired when asked for by
  a timeout''}:
%
$\mathtt{AG}(\PortAsk \rightarrow \mathtt{AF}\ \PortReset)$.
%
The satisfaction of the safety property could be established by
applying symbolic model checking techniques.  However, in this
example, it is obvious by inspection of the open automaton.  The
satisfaction of the liveness property relies on the above observation
that in the state $\ostate{s2,t0}$ only the reset transition involves
$\NameCtrl$.  Therefore, under resonable scheduling assumptions,
$\PortReset$ will always be fired.

%% \TODO{it would be good to express here maybe 2 properties of the
%%   architecture, maybe one taken from the original study, and another
%%   one involving the timer data (number of ticks ?)}

@@

In this paper, we presented an algorithm for generating an open
automaton representing the semantics of an open pNet.
We have also presented our implementation, which comprises two main parts. Firstly, we
compute all the possible open transitions. \noteSB{I do not understand this sentence. Suggestion: The actions of pLTS and hole behaviours are composed by the algoritm during the generation of predicates needed to match...(?)}\hl{The actions in pLTS and
hole behaviors are composed in the algorithm, while the generation of
predicate need to match the combinations of the subnets with the
synchronization vectors.}  Some of the open transitions obtained at this
intermediate step, constructed in a structural manner from all
possible combinations of possible logical predicates, may
contain predicates which do not represent any possible concrete
instantiations. 
Secondly, to get rid of these useless transitions, we use the SMT solver Z3 for checking the
satisfiability of the predicate in each open transition. 
To this end, we encode into Z3 the representations of the action algebra and
the predicates before submitting them to the
Z3 solver. We implemented our algorithm in the VerCors platform and 
tested the algorithm on available examples, such as encoding expressions from various process algebras. Our case-studies show that our encoding
successfully identifies all unsatisfiable open transitions and that
the resulting automata correctly reflect the expected 
movements of the encoded process expressions.

Among the questions that have arised during this work, an important one is the
efficiency of Z3 for our needs. Z3 is famous for being very fast in
solving very large sets of assertions.%% , which could be important for
%% us
However, we encountered problems during the implementation of the
checking method using Z3. For example, if the $\mathit{functions}$ and the
$\mathit{datatypes}$ part of the algebra presentation that we submit to the solver
are independent from each other, then checking the $\mathit{assertions}$ is 
simple. However, $\mathit{functions}$ on recursive $\mathit{datatypes}$ make it
more complex. Special rules might be defined by the user for the
induction. Also, we need more evidence that dealing with more complex
action algebras, which would involve axiomatizing their data structures
into Z3 theories, will indeed allow us to decide validity of
predicates.
We may also try other automatic theorem provers, depending on the structure of the proofs we need.

@@

Naturally, our next goals after the generation of the open automata
will be to model-check, and to check equivalence of  pNets. While
model-checking open automata (with a suitable logic including
predicates on data) seems easy to define, equivalence checking is more challenging.
In a previous paper~\hl{\cite{x}}, we
have already found the FH-bisimulation, inspired by~\cite{deSimone85}, to be a
suitable definition. However, the usual approach of computing bisimulation by partition
refinement does not work easily with symbolic transitions. In a first
step we may want instead to define and implement an algorithm for
checking that a given equivalence relation between open automata 
states is indeed a FH-Bisimulation. Finding automatically a
bisimulation relation or the most general bisimulation will be more
challenging. It appears also that when comparing different pNets with
different structure, strong bisimulation will be too strong a notion,
and that we will have to define proper notions of weak equivalence or
refinements.

@@

\begin{table}\caption{\label{table:CCSalgebra}\leftline{Additional
      sorts and operators for the CCS algebra}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Auxiliary Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Channel, Data    			&
                $\emptyset$&
                $\emptyset$
                \\\specialrule{0em}{1pt}{1pt} 
		Action 			&  $\texttt{Emit(c,v)},\texttt{Receive(c,v)},\texttt{Tau}$ &
		\\\hline
	\end{tabular}
\end{table}

@@

		&&       		\\\specialrule{0em}{1pt}{1pt}
		\multicolumn{2}{l}{Auxiliary operators: }
